---
title: "R Notebook GM"
output:
  html_document:
    df_print: paged
---

```{r}
library(ggplot2)
library(corrplot)
library(dplyr)
library(GGally)
library(pheatmap)
library(tibble)
library(maps)

```

## Importing dataset from Github 

```{r}
url_df='https://raw.githubusercontent.com/liubas3171/SL_group_project/main/Unsupervised%20Learning/data/Life%20Expectancy%20Data.csv'

data <- read.csv(url_df)



```

## Data exploration of the dataset


```{r}
na_count <- sum(is.na(data))
print(paste('Absolute value of null values:',na_count))
data <- na.omit(data)
```

```{r}
## Value counts for the Region field
```{r}
table(data$Region)
```
## Summary of continuous variables

```{r}
summary(data)
```

## Preparing the dataset for PCA. By fixing the "Country" column as the index of the dataset, extract all continuous variable columns.
```{r}
cols <- c('Region','Population_mln', 'Economy_status_Developed', 'Year','Total.expenditure','Income.composition.of.resources','percentage.expenditure','GDP_per_capita','Schooling','Life_expectancy')
df_pca <- data  
df_pca <- df_pca[, !colnames(df_pca) %in% cols]
df_pca_agg <- aggregate(. ~ Country, df_pca, mean, na.rm = TRUE)
rownames(df_pca_agg) <- df_pca_agg[, "Country"]
df_pca_agg <- df_pca_agg[, -which(names(df_pca_agg) == "Country")]  
df_pca_agg
```
### PCA
```{r}
pca_res = prcomp(df_pca_agg, scale=TRUE)
pca_res
```

## Result of pca : PC1 variability explains the 42,5% of totale variance of data
```{r}
summary(pca_res)
```


```{r}
biplot(pca_res, scale=0, cex=0.55)
```
```{r}
screeplot(princomp(data[], cor=T))
```
```{r}
temp_v <- pca_res$sdev^2
temp_v <- temp_v / sum(temp_v)
#temp_v
plot(temp_v, xlab="Principal Component", ylab="Proportion of Variance Explained", ylim=c(0,1),type='b')
```
```{r}
plot(cumsum(temp_v), xlab="Principal Component", ylab="Cumulative Proportion of Variance Explained", ylim=c(0,1),type='b')
```

```{r}
rho <- cor(data[pca_res_colnames])
autoval <- eigen(rho)$values
plot(autoval, type="b", main="Scree Diagram", xlab="Number of Component", ylab="Eigenvalues")
abline(h=1, lwd=3, col="red")
```
```{r}

```

## Сlustering

### GDP_per_capita & Infant_deaths
```{r}
## Clusterization on life expectancy and percentage expenditure

cl_gdp_id <- dist(scale(data[,c('GDP_per_capita', 'Infant_deaths')]), method = 'euclidean')

# Decide how many clusters to look at
n_clusters <- 10

# Initialize total within sum of squares error: wss
wss <- numeric(n_clusters)

set.seed(123)

# Look over 1 to n possible clusters
for (i in 1:n_clusters) {
  # Fit the model: km.out
  km.out <- kmeans(cl_gdp_id, centers = i, nstart = 20)
  # Save the within cluster sum of squares
  wss[i] <- km.out$tot.withinss
}

# Produce a scree plot
wss_df <- tibble(clusters = 1:n_clusters, wss = wss)
 
scree_plot <- ggplot(wss_df, aes(x = clusters, y = wss, group = 1)) +
    geom_point(size = 4)+
    geom_line() +
    scale_x_continuous(breaks = c(2, 4, 6, 8, 10)) +
    xlab('Number of clusters')

```

# Visualize cluster 
```{r}
k <- 3
set.seed(123)
# Build model with k clusters: km.out
km.out <- kmeans(cl_gdp_id, centers = k, nstart = 20)



data$cluster_id <- factor(km.out$cluster)


ggplot(data, aes(x = GDP_per_capita, y = Infant_deaths, color = factor(cluster_id))) +
    geom_point(alpha = 0.25) +
    xlab("GDP_per_capita") +
    ylab("Infant_deaths") +
    ggtitle("GDP_per_capita & Infant_deaths") +
    theme(plot.title = element_text(hjust = 0.5))


```
```{r}
# Top 5 undevelop countries (cluster 1) for infatn_deaths in relation to


mean_data <- aggregate(cbind(Infant_deaths, GDP_per_capita,cluster_id) ~ Country, data = data, FUN = mean)


top_countries <- subset(mean_data, cluster_id == 1)


top_countries <- top_countries[order(top_countries$Infant_deaths, decreasing = TRUE), ]


top_countries <- head(top_countries, 5)


top_countries[, c('Country', 'Infant_deaths', 'GDP_per_capita')]
```
## GDP & Alchool 
```{r}
cl_gdp_alchool <- dist(scale(data[,c('GDP_per_capita', 'Alcohol_consumption')]), method = 'euclidean')

# Decide how many clusters to look at
n_clusters <- 10

# Initialize total within sum of squares error: wss
wss <- numeric(n_clusters)

set.seed(123)

# Look over 1 to n possible clusters
for (i in 1:n_clusters) {
  # Fit the model: km.out
  km.out <- kmeans(cl_gdp_alchool, centers = i, nstart = 20)
  # Save the within cluster sum of squares
  wss[i] <- km.out$tot.withinss
}

# Produce a scree plot
wss_df <- tibble(clusters = 1:n_clusters, wss = wss)
 
scree_plot <- ggplot(wss_df, aes(x = clusters, y = wss, group = 1)) +
    geom_point(size = 4)+
    geom_line() +
    scale_x_continuous(breaks = c(2, 4, 6, 8, 10)) +
    xlab('Number of clusters')
```

```{r}
k <- 3
set.seed(123)
# Build model with k clusters: km.out
km.out <- kmeans(cl_gdp_alchool, centers = k, nstart = 20)



data$cluster_id <- factor(km.out$cluster)


ggplot(data, aes(x = GDP_per_capita, y = Alcohol_consumption, color = factor(cluster_id))) +
    geom_point(alpha = 0.25) +
    xlab("GDP_per_capita") +
    ylab("Alcohol_consumption") +
    ggtitle("GDP_per_capita & Alcohol_consumption") +
    theme(plot.title = element_text(hjust = 0.5))
```
```{r}
# Top 5  countries  for Aalchool consumption in relation to gdp


mean_data <- aggregate(cbind(Alcohol_consumption, GDP_per_capita,cluster_id) ~ Country, data = data, FUN = mean)


top_countries <- subset(mean_data, cluster_id == 1)


top_countries <- top_countries[order(top_countries$Alcohol_consumption, decreasing = TRUE), ]


top_countries <- head(top_countries, 5)


top_countries[, c('Country', 'Alcohol_consumption', 'GDP_per_capita')]
```






