---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

```{r}
library(ggplot2)
library(corrplot)
```



```{r}
data <- read.csv("../LifeExpectancyData.csv")



```

## General aggregations


```{r}
na_count <- sum(is.na(data))
na_count
data <- na.omit(data)
sum(is.na(data))
```

```{r}
head(table(data$ï..Country))
length(unique(data$ï..Country))
```

```{r}
table(data$Region)
```

```{r}
sum(data$Economy_status_Developed)
```




## Summary of continuous variables


```{r}
summary(data)
```

```{r}
categorical_vars <- c('ï..Country', 'Region', 'Economy_status_Developed', 'Year')
my_colnames <- colnames(data)
my_colnames <- my_colnames[!my_colnames %in% categorical_vars] 
my_colnames
```
## Some analysis

```{r}
curr_data <- data[data$Year == 2015,]
curr_data <- curr_data[curr_data$Region == 'European Union',]
```





### PCA

NORMALIZATION OF VARS - scale=true should do so
WHAT TO DO WITH YEARS -?
GROUP VARIABLES BY HELTH INDICATIONS, ECONOMIC, ETC
play with countries, play with variables

prcomp() - to do PCA

```{r}
rownames(curr_data) <- curr_data$ï..Country
curr_data
```
```{r}
cor_matrix <- cor(curr_data[my_colnames])
corrplot(cor_matrix, method = "color", tl.cex=0.6)
```
```{r}
help(corrplot)
```


```{r}
#curr_data[my_colnames]
pca_colnames = c(
       "Life_expectancy",
       #"Adult_mortality",
       "Infant_deaths",
       #"Hepatitis_B",
       "BMI",
       #"Measles",
       #"Polio",
       "Alcohol_consumption"
       )
```



```{r}
pca_res = prcomp(curr_data[pca_colnames], scale=TRUE)
pca_res
```

```{r}
pca_res$x
```


```{r}
biplot(pca_res, scale=0, cex=0.6)
```
```{r}
screeplot(princomp(curr_data[pca_colnames], cor=T))
```
```{r}
temp_v <- pca_res$sdev^2
temp_v <- temp_v / sum(temp_v)
#temp_v
plot(temp_v, xlab="Principal Component", ylab="Proportion of Variance Explained", ylim=c(0,1),type='b')
```
```{r}
plot(cumsum(temp_v), xlab="Principal Component", ylab="Cumulative Proportion of Variance Explained", ylim=c(0,1),type='b')
```

```{r}
rho <- cor(curr_data[pca_colnames])
autoval <- eigen(rho)$values
plot(autoval, type="b", main="Scree Diagram", xlab="Number of Component", ylab="Eigenvalues")
abline(h=1, lwd=3, col="red")
```



```{r}

```









## Сlustering

### Hierarchical clustering


#### Life expectancy
```{r}
res_dist <- dist(curr_data["Life_expectancy"], method = "euclidean")
hclust_res <- hclust(res_dist, method="complete")
plot(hclust_res)
```

#### Alcohol consumption
```{r}
res_dist <- dist(curr_data["Alcohol_consumption"], method = "euclidean")
hclust_res <- hclust(res_dist, method="complete")
plot(hclust_res)
```
#### Schooling
```{r}
res_dist <- dist(curr_data["Schooling"], method = "euclidean")
hclust_res <- hclust(res_dist, method="complete")
plot(hclust_res)
```


#### Health indicators
```{r}


health_cols = c("Hepatitis_B", "Measles", "Polio", "Diphtheria", "Incidents_HIV")
normalized_data <- scale(curr_data[health_cols])
res_dist <- dist(normalized_data, method = "euclidean")
hclust_res <- hclust(res_dist, method="complete")
plot(hclust_res)
```
```{r}
curr_data
```



```{r}
curr_data
```




```{r}
curr_data["Life_expectancy"]
```



### K-means

```{r}

curr_data
```

```{r}
summary(curr_data)
```


```{r}
curr_rows <- c('Life_expectancy', 'Adult_mortality') #, 'Infant_deaths', 'Alcohol_consumption', 'BMI')
curr_data <- curr_data[curr_rows]
```


```{r}
kmeans_res <- kmeans(curr_data, centers=4)
#kmeans_res
```

```{r}

plot(curr_data[, 1], curr_data[, 2], col = kmeans_res$cluster, pch = 20)
#points(kmeans_res$centers[, 1], kmeans_res$centers[, 2], col = 1:k, pch = 3, cex = 2)

```


```{r}
k = 5
kmeans_res_pca = kmeans(pca_res$x, centers=k)

```

```{r}
plot(pca_res$x[, 1], pca_res$x[, 2], col = kmeans_res_pca$cluster, pch = 20)
points(kmeans_res_pca$centers[, 1], kmeans_res_pca$centers[, 2], col = 1:k, pch = 3, cex = 2)
```

```{r}
pca_res$x[, 1]
```








